---
title: "R Notebook"
output: html_notebook
---

#load all relevant packages
```{r}
library(dplyr)
library(tidyverse)
library(zoo)
library(pracma)
library(lubridate)
library(tibble)
library(ggplot2)
library(targets)
library(tseries)
library(urca)
library(forecast)
```

#Load the cleaned datasets, under all timescales
```{r}
dfFull <- tar_read(combined_15min_data)
df_15min <- tar_read(combined_15min_data) %>% select(date, height, station_id)
df_hourly <- tar_read(combined_hourly_data)
df_daily <- tar_read(combined_daily_data)
df_weekly <- tar_read(combined_weekly_data)
```

#Counting missing values/outliers after data cleaning
```{r}
NACountfunc <- function(df) {
  NACountDF <- data.frame(stations = riverStations,
                          MissingNA = numeric(33),
                          OutlierNA = numeric(33),
                          percentage = numeric(33))
  
  for (i in seq_along(riverStations)) {
    dfSingle <- df %>% filter(station_id == riverStations[i])
    NACountDF$MissingNA[i] <- sum(is.na(dfSingle$value))
    NACountDF$OutlierNA[i] <- sum(is.na(dfSingle$newValue)) - NACountDF$MissingNA[i]
    NACountDF$percentage[i] <- round(100 * sum(is.na(dfSingle$newValue))/dim(dfSingle)[1], 2)
  }
  
  return(NACountDF)
}

NAcounts <- NACountfunc(dfFull)
NAcounts
```

#Summary statistics (Full table in appendices)
```{r}
GetSummaryStats <- function(dfAll) {
  summaryStatsDF <- data.frame(stations = riverStations,
                               Min = numeric(33),
                               firstQu = numeric(33),
                               Median = numeric(33),
                               Mean = numeric(33),
                               thirdQu = numeric(33),
                               Max = numeric(33))
  
  for (i in seq_along(riverStations)) {
    dfStation <- dfAll %>% filter(station_id == riverStations[i])
    
    summaryStatsDF$Min[i] <- as.numeric(summary(dfStation$height)[1])
    summaryStatsDF$firstQu[i] <- as.numeric(summary(dfStation$height)[2])
    summaryStatsDF$Median[i] <- as.numeric(summary(dfStation$height)[3])
    summaryStatsDF$Mean[i] <- as.numeric(summary(dfStation$height)[4])
    summaryStatsDF$thirdQu[i] <- as.numeric(summary(dfStation$height)[5])
    summaryStatsDF$Max[i] <- as.numeric(summary(dfStation$height)[6])
  }
  
  return(summaryStatsDF)
}

GetSummaryStats(dfFull)
```



Corresponding plots for outliers/missing values (numbering conformed with the dissertation project)
#Figure 3
```{r}
stationPlotfunction <- function(id, start, end) {
  df <- dfFull %>% filter(station_id == as.character(id)) %>%
                   filter(date >= start & date <= end)
  
  ggplot(df, aes(x = date)) +
  geom_line(aes(y = value), size = 0.7, color = "blue") +
  labs(x = "Date", y = "River height (m)") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10, face = "bold")) +
  theme(axis.text.y = element_text(size = 10, face = "bold")) +
  theme(panel.grid.minor= element_blank(),
        axis.title.x = element_text(size = 12, face = "bold"), 
        axis.title.y = element_text(size = 12, face = "bold"))
}

stationPlotfunction(7038, as.POSIXct("2019-07-22 00:00:00", tz = "UTC"),
                          as.POSIXct("2019-07-24 00:00:00", tz = "UTC"))
stationPlotfunction(7057, as.POSIXct("2024-02-24 00:00:00", tz = "UTC"),
                          as.POSIXct("2024-02-26 00:00:00", tz = "UTC"))
```

#Figure 4
```{r}
stationPlotfunction(7048, as.POSIXct("2019-08-08 00:00:00", tz = "UTC"),
                          as.POSIXct("2019-08-10 00:00:00", tz = "UTC"))

stationPlotfunction(7055, as.POSIXct("2021-07-01 00:00:00", tz = "UTC"),
                          as.POSIXct("2021-07-09 00:00:00", tz = "UTC"))
```

#Figure 5
```{r}
OutlierPlotfunction <- function(id, start, end) {
  df <- dfFull %>% filter(station_id == as.character(id)) %>%
                   filter(date >= start & date <= end)
  
  ggplot(df, aes(x = date)) +
  geom_line(aes(y = value), size = 0.7, color = "red") +
  geom_line(aes(y = newValue), size = 0.7, color = "blue") +
  labs(x = "Date", y = "River height (m)") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10, face = "bold")) +
  theme(axis.text.y = element_text(size = 10, face = "bold")) +
  theme(panel.grid.minor= element_blank(),
        axis.title.x = element_text(size = 12, face = "bold"), 
        axis.title.y = element_text(size = 12, face = "bold"))
}

OutlierPlotfunction(7038, as.POSIXct("2019-07-22 00:00:00", tz = "UTC"),
                          as.POSIXct("2019-07-24 00:00:00", tz = "UTC"))

OutlierPlotfunction(7057, as.POSIXct("2024-02-24 00:00:00", tz = "UTC"),
                          as.POSIXct("2024-02-26 00:00:00", tz = "UTC"))
```



Chapter 4: Results
#River stations
```{r}
riverThames <- c("7027", "7034", "7035", "7036", "7037", "7038", "7046", "7048", "7049", "7055", "7056", "7057", "7072", "7405", "7073", "7083", "7085", "7086")
riverOck <- c("7402", "7078", "7077")
riverCherwell <- c("7059", "1414", "7063", "7064", "7066", "7071")
streamsBrooks <- c("7039", "7001", "7075", "7074", "7076", "7003")
riverStations <- c(riverThames, riverOck, riverCherwell, streamsBrooks)
```

4.1: Mechanistic models
#Plots for illustrative examples: Figure 7 (Stations 7034 and 7035)
```{r}
df7035 <- df_15min %>% filter(station_id == "7034")
df7036 <- df_15min %>% filter(station_id == "7035")
df_combined <- data.frame(
  Index = rep(df7035$date[25:73], 2),
  Height = c(df7035$height[25:73], df7036$height[25:73]),  
  Stations = rep(c("7034", "7035"), each = 49)  
)

ggplot(df_combined, aes(x = Index, y = Height, color = Stations)) +
  geom_line(size = 0.8) + 
  geom_vline(xintercept = df7035$date[35], linetype = "dashed", color = "black", size = 0.5) +
  geom_vline(xintercept = df7035$date[45], linetype = "dashed", color = "black", size = 0.5) +
  geom_vline(xintercept = df7035$date[55], linetype = "dashed", color = "green", size = 0.5) +
  geom_vline(xintercept = df7035$date[70], linetype = "dashed", color = "green", size = 0.5) +
  labs(x = "time",
       y = "River height (m)") +
  theme_minimal() +
  theme(axis.text.x = element_text(size = 10, face = "bold")) +
  theme(axis.text.y = element_text(size = 10, face = "bold")) +
  theme(panel.grid.major= element_blank(),
        axis.title.x = element_text(size = 12, face = "bold"), 
        axis.title.y = element_text(size = 12, face = "bold"))
```

#Plot function for adjusted Rsquare & AIC against tau: like Figures 8, 9, 11, 12, 22, 23
```{r}
plotFunctionAcrossTaus <- function(tausVec, infoList) {
  plotData <- data.frame(tausVec, infoList$AIC, infoList$Rsquare)
  
  pic1 <- ggplot(plotData, aes(x = tausVec, y = infoList$AIC)) +
            geom_line(color = "blue") +
            geom_point(color = "blue") +
            labs(title = expression("AIC vs." ~ bold(tau)), x = expression(bold(tau)), y = "AIC") +
            theme_minimal() + 
            theme(axis.text.x = element_text(size = 10, face = "bold")) +
            theme(axis.text.y = element_text(size = 10, face = "bold")) +
            theme(
                  axis.title.x = element_text(size = 13, face = "bold"), 
                  axis.title.y = element_text(size = 13, face = "bold"))

  pic2 <- ggplot(plotData, aes(x = tausVec, y = infoList$Rsquare)) +
            geom_line(color = "skyblue") +
            geom_point(color = "skyblue") +
            labs(title = expression("Adjusted" ~ R^2 ~ "vs." ~ bold(tau)), x = expression(bold(tau)), y = expression("Adjusted" ~ R^2)) +
            theme_minimal()+
            theme(axis.text.x = element_text(size = 10, face = "bold")) +
            theme(axis.text.y = element_text(size = 10, face = "bold")) +
            theme(
                  axis.title.x = element_text(size = 13, face = "bold"), 
                  axis.title.y = element_text(size = 13, face = "bold"))
  
  print(pic1)
  print(pic2)
}
```


Fitting simple linear regression models
#Retrieve training data (with specific tau)
```{r}
retrieveTrainingDf <- function(id, Fulldf, tau) {
  upstreamPosition <- NA
  stationPosition <- which(riverStations == as.character(id))
  
  if (stationPosition == 1) {
    df <- Fulldf %>% filter(station_id == id) %>%
                   mutate(heightU = 0) %>% 
                   select(date, heightU, height, station_id)
  } else if (adjMatrix1[stationPosition-1, stationPosition] == 1) {
    upstreamPosition <- stationPosition - 1
    dfUpstream <- Fulldf %>% filter(station_id == riverStations[upstreamPosition])
    df <- Fulldf %>% filter(station_id == id) %>%
                     mutate(heightU = dfUpstream$height) %>% 
                     select(date, heightU, height, station_id)
  } else {
    df <- Fulldf %>% filter(station_id == id) %>%
                   mutate(heightU = 0) %>% 
                   select(date, heightU, height, station_id)
  }
  
  if (tau < nrow(df)) {
    df <- df %>% mutate(heightU = c(rep(NA, tau), heightU[1:(nrow(df) - tau)]))
    df <- df %>% filter(!is.na(heightU))
  } else {
    warning("tau too large!!")
  }
  return(df)
}
```

#Obtain corresponding adjusted Rsquares of models with optimal tau
```{r}
LMtauList15min <- list(1, seq(1,5), seq(8,12), seq(3,7), seq(5,10),
                       seq(20,30), seq(23,27), seq(1,5), seq(1,5), seq(1,5),
                       seq(245,250), seq(1,5), seq(5,10), seq(2,6), seq(6,10),
                       seq(1,5), seq(15,20), seq(1,5), 1, seq(65,70), 
                       seq(1,5), 1, seq(25,30), seq(130,135), seq(32,37),
                       seq(1,5), seq(16,21), 1,1,1,1,1,1)

LMtauListHourly <- rep(list(seq(1, 5)), 33)
LMtauListHourly[[11]] <- seq(60,65)
LMtauListHourly[[20]] <- seq(15,20)
LMtauListHourly[[23]] <- seq(5,10)
LMtauListHourly[[24]] <- seq(30,35)
LMtauListHourly[[25]] <- seq(8,13)
LMtauListHourly[[27]] <- seq(3,8)
  
LMtauListDaily <- rep(list(seq(1, 5)), 33)
LMtauListWeekly <- rep(list(seq(1, 5)), 33)

retrieveOptRsquare <- function(df, tausList) {
  Rsquares <- vector("numeric", 33)
  
  for (i in seq_along(riverStations)) {
    for (j in seq_along(tausList[[i]])) {
      trainingDf <- retrieveTrainingDf(riverStations[i], df, tausList[[i]][j])
      model <- lm(height ~ 1 + heightU, data = trainingDf)
      if (summary(model)$adj.r.squared > Rsquares[i]) {
        Rsquares[i] <- summary(model)$adj.r.squared
      }
    }
  }
  
  return(Rsquares[Rsquares != 0])
}

Rsquare15min <- retrieveOptRsquare(df_15min, LMtauList15min)
RsquareHourly <- retrieveOptRsquare(df_hourly, LMtauListHourly)
RsquareDaily <- retrieveOptRsquare(df_daily, LMtauListDaily)
RsquareWeekly <- retrieveOptRsquare(df_weekly, LMtauListWeekly)
```

#Return optimal tau, and corresponding time in the unit of minutes
```{r}
retrieveOptTaus <- function(df, tausList) {
  Taus <- vector("integer", 33)
  
  for (i in seq_along(riverStations)) {
    Rsquares <- vector("numeric", length(tausList[[i]]))
    for (j in seq_along(tausList[[i]])) {
      trainingDf <- retrieveTrainingDf(riverStations[i], df, tausList[[i]][j])
      model <- lm(height ~ 1 + heightU, data = trainingDf)
      Rsquares[j] <- summary(model)$adj.r.squared
    }
    Taus[i] <- tausList[[i]][which.max(Rsquares)]
  }
  
  return(Taus)
}

#removing those without an adjacent upstream station
removedIndices <- c(1,19,22,28,29,30,31,32,33)
optTau15min <- retrieveOptTaus(df_15min, LMtauList15min)
optTauHourly <- retrieveOptTaus(df_hourly, LMtauListHourly)
optTauDaily <- retrieveOptTaus(df_daily, LMtauListDaily)
optTauWeekly <- retrieveOptTaus(df_weekly, LMtauListWeekly)

#convert optimal tau values into corresponding time, in the unit of minutes
TauDF <- data.frame(stations = riverStations[-removedIndices],
                    mins = 15*optTau15min[-removedIndices],
                    Hourly = 60*optTauHourly[-removedIndices],
                    Daily = 1440 *optTauDaily[-removedIndices],
                    Weekly = 10080*optTauWeekly[-removedIndices])
```

#Point range plots under all timescales: Figure 10
```{r}
ggplot(pointRangeDF, aes(x = timeScale, y = median)) +
  theme_minimal() +
  geom_pointrange(aes(ymin = lower, ymax = upper), size = 0.5, color = "skyblue") +
  xlab("time scale") +
  ylab(bquote(bold("Adjusted" ~ R^2))) +
  geom_text(aes(label = sprintf("%.4f", median)), vjust = -1.5, color = "black") +
  geom_text(aes(y = lower, label = sprintf("%.4f", lower)), vjust = 1.5, color = "red") +
  geom_text(aes(y = upper, label = sprintf("%.4f", upper)), vjust = -0.5, color = "blue") +
  theme(axis.text.x = element_text(size = 10, face = "bold")) +
  theme(axis.text.y = element_text(size = 10, face = "bold")) +
  theme(panel.grid.minor= element_blank(),
        axis.title.x = element_text(size = 14, face = "bold"), 
        axis.title.y = element_text(size = 14, face = "bold"))
```




4.2 Autoregressive models
#Retrieve training datasets for AR(1) with upstream term (for a specific tau)
```{r}
retrieveTrainingDf2 <- function(id, Fulldf, tau) {
  upstreamPosition <- NA
  stationPosition <- which(riverStations == as.character(id))
  
  if (stationPosition == 1) {
    df <- Fulldf %>% filter(station_id == id) %>%
                     mutate(heightU = 0) %>% 
                     select(date, heightU, height, station_id)
  } else if (stationPosition != 1 & adjMatrix1[stationPosition-1, stationPosition] == 1) {
    upstreamPosition <- stationPosition - 1
    dfUpstream <- Fulldf %>% filter(station_id == riverStations[upstreamPosition])
    df <- Fulldf %>% filter(station_id == id) %>%
                     mutate(heightU = dfUpstream$height) %>% 
                     select(date, heightU, height, station_id)
  } else {
    df <- Fulldf %>% filter(station_id == id) %>%
                   mutate(heightU = 0) %>% 
                   select(date, heightU, height, station_id)
  }
  
  if (tau < nrow(df)) {
    df <- df %>% mutate(heightU = c(rep(NA, tau), heightU[1:(nrow(df) - tau)]))
  } else {
    warning("tau too large!!")
  }

  df <- df %>% mutate(heightAR = lag(height, 1)) %>%
               filter(!is.na(heightU)) %>%
               filter(!is.na(heightAR)) %>%
               select(date, height, heightAR, heightU, station_id)
  
  return(df)
}
```

#Retrieve the adjusted Rsquare for AR(1) and AR(1) with upstream term (optimal tau)
```{r}
#candidates of optimal tau (only small tau values)
ARtauList15min <- rep(list(seq(1, 5)), 33)
ARtauListHourly <- rep(list(seq(1, 5)), 33)
ARtauListDaily <- rep(list(seq(1, 5)), 33)
ARtauListWeekly <- rep(list(seq(1, 5)), 33)

retrieveOptRsquareAR <- function(df, tausList) {
  Rsquares <- vector("numeric", 33)
  RsquaresARonly <- vector("numeric", 33)
  
  for (i in seq_along(riverStations)) {
    dfARonly <- df %>% filter(station_id == riverStations[i]) %>%
                       mutate(heightAR = lag(height, 1)) %>%
                       select(date, heightAR, height)
    model1 <- lm(height ~ 1 + heightAR, data = dfARonly)
    RsquaresARonly[i] <- summary(model1)$adj.r.squared
    
    for (j in seq_along(tausList[[i]])) {
      trainingDf <- retrieveTrainingDf2(riverStations[i], df, tausList[[i]][j])
      model2 <- lm(height ~ 1 + heightAR + heightU, data = trainingDf)
      
      if (summary(model2)$adj.r.squared > Rsquares[i]) {
        Rsquares[i] <- summary(model2)$adj.r.squared
      }
    }
  }
  
  return(list(RsquaresARonly, Rsquares))
}

Rsquare15minAR <- lapply(retrieveOptRsquareAR(df_15min, ARtauList15min), function(vec) vec[-removedIndices])
RsquareHourlyAR <- lapply(retrieveOptRsquareAR(df_hourly, ARtauListHourly), function(vec) vec[-removedIndices])
RsquareDailyAR <- lapply(retrieveOptRsquareAR(df_daily, ARtauListDaily), function(vec) vec[-removedIndices])
RsquareWeeklyAR <- lapply(retrieveOptRsquareAR(df_weekly, ARtauListWeekly), function(vec) vec[-removedIndices])
```

#Point range plot: Figure 13
```{r}
ARpointRangeDF <- data.frame(
                           colour = rep(c("AR only", "with upstream term"), each = 4),
                           timeScale = factor(rep(c("15min", "hourly", "daily", "weekly"), 2), levels = c("15min", "hourly", "daily", "weekly")), 
                           median = c(median(Rsquare15minAR[[1]]), median(RsquareHourlyAR[[1]]), median(RsquareDailyAR[[1]]), median(RsquareWeeklyAR[[1]]),median(Rsquare15minAR[[2]]), median(RsquareHourlyAR[[2]]), median(RsquareDailyAR[[2]]), median(RsquareWeeklyAR[[2]])),
                           lower = c(quantile(Rsquare15minAR[[1]], 0.025), quantile(RsquareHourlyAR[[1]], 0.025), quantile(RsquareDailyAR[[1]], 0.025), quantile(RsquareWeeklyAR[[1]], 0.025), quantile(Rsquare15minAR[[2]], 0.025), quantile(RsquareHourlyAR[[2]], 0.025), quantile(RsquareDailyAR[[2]], 0.025), quantile(RsquareWeeklyAR[[2]], 0.025)),
                           upper = c(quantile(Rsquare15minAR[[1]], 0.975), quantile(RsquareHourlyAR[[1]], 0.975), quantile(RsquareDailyAR[[1]], 0.975), quantile(RsquareWeeklyAR[[1]], 0.975), quantile(Rsquare15minAR[[2]], 0.975), quantile(RsquareHourlyAR[[2]], 0.975), quantile(RsquareDailyAR[[2]], 0.975), quantile(RsquareWeeklyAR[[2]], 0.975)))

ggplot(ARpointRangeDF, aes(x = timeScale, y = median, color = colour)) +
  theme_minimal() +
  geom_pointrange(aes(ymin = lower, ymax = upper), position = position_dodge(width = 1)) +
  geom_text(aes(label = sprintf("%.4f", median)), position = position_dodge(width = 1), vjust = -2, hjust = 0.5, size = 3.4) +
  geom_text(aes(y = lower, label = sprintf("%.4f", lower)), position = position_dodge(width = 1), vjust = 1, hjust = 0.5, size = 3.4) +
  geom_text(aes(y = upper, label = sprintf("%.4f", upper)), position = position_dodge(width = 1), vjust = -0.4, hjust = 0.5, size = 3.4) +
  xlab("timescale") +
  ylab(bquote(bold("Adjusted" ~ R^2))) +
  scale_color_manual(values = c("AR only" = "skyblue", "with upstream term" = "black"))+
  labs(color = "models") +
  theme(axis.text.x = element_text(size = 10, face = "bold")) +
  theme(axis.text.y = element_text(size = 10, face = "bold")) +
  theme(panel.grid.minor= element_blank(),
        axis.title.x = element_text(size = 12, face = "bold"), 
        axis.title.y = element_text(size = 12, face = "bold"))
```

#Fitted models example plots: Figure 14  and 15
```{r}
stationARplot <- function(id, dfAll, interval) {
  dfAR <- dfAll %>% filter(station_id == as.character(id)) %>%
                    mutate(heightAR = lag(height, 1)) %>%
                    select(date, heightAR, height)
  modelAR <- lm(height ~ 1 + heightAR, data = dfAR)
  
  dfAR <- dfAR[-1,]
  observations <- dfAR$height
  fitted <- modelAR$fitted.values
  
  dfPlot <- data.frame(time = dfAR$date[interval],
                       Observations = observations[interval],
                       Fitted = fitted[interval]
                       )
  
  ggplot(dfPlot, aes(x = time)) +
  geom_line(aes(y = Observations, color = "Observations"), size = 1) +
  geom_line(aes(y = Fitted, color = "Fitted Values"), size = 1) +
  labs(x = "time",
       y = "River height (m)",
       color = "Data type") +
  theme_minimal() +
      theme(axis.text.x = element_text(size = 10, face = "bold")) +
      theme(axis.text.y = element_text(size = 10, face = "bold")) +
      theme(panel.grid.minor= element_blank(),
            axis.title.x = element_text(size = 14, face = "bold"), 
            axis.title.y = element_text(size = 14, face = "bold"))
}

stationARplot(7064, df_15min, 100:180)
stationARplot(7064, df_hourly, 100:180)
stationARplot(7064, df_daily, 100:180)
stationARplot(7064, df_weekly, 100:180)

stationARplot(7034, df_15min, 100:180)
stationARplot(7034, df_hourly, 100:180)
stationARplot(7034, df_daily, 100:180)
stationARplot(7034, df_weekly, 100:180)
```



4.3 ARMA/ARIMA/SARIMA models
#Stationary test (ADF and KPSS)
```{r}
retrieveStationarityInfo <- function(dfALL) {
  stationarityCheckdf <- data.frame(stations = riverStations,
                                  ADF = numeric(33),
                                  KPSS = numeric(33),
                                  status = numeric(33))
  
  for (i in seq_along(riverStations)) {
    df <- dfALL %>% filter(station_id == riverStations[i])
    stationarityCheckdf$ADF[i] <- adf.test(df$height)$p.value
    stationarityCheckdf$KPSS[i] <- kpss.test(df$height)$p.value
    
    if (adf.test(df$height)$p.value < 0.05 & kpss.test(df$height)$p.value >0.05) {
      stationarityCheckdf$status[i] <- 1
    } else {
      stationarityCheckdf$status[i] <- 0
    }
  }
  
  return(stationarityCheckdf)
}

dfStationarityDaily <- retrieveStationarityInfo(df_daily)
dfStationarityweekly <- retrieveStationarityInfo(df_weekly)
```

#Fit ARMA/ARIMA models
#Comparing AIC values for different models
```{r}
ARModelsAIC1 <- function(dfAll) {
  statsDF <- data.frame(stations = riverStations,
                              ARMA = numeric(33),
                              ARIMA = numeric(33)
                        )
  
  for (i in seq_along(riverStations)) {
    dfModel <- dfAll %>% filter(station_id == riverStations[i]) 
  
    model1 <- auto.arima(dfModel$height, d = 0, seasonal = FALSE, stepwise = TRUE, ic = "aic")
    model2 <- auto.arima(dfModel$height, seasonal = FALSE, stepwise = TRUE, ic = "aic")
    
    statsDF$ARMA[i] <- model1$aic
    statsDF$ARIMA[i] <- model2$aic
  }
  
  statsDF$diff <- statsDF$ARMA - statsDF$ARIMA
  return(statsDF)
}

#for daily data
dfCompareDaily <- ARModelsAIC1(df_daily)
#for weekly data
dfCompareWeekly <- ARModelsAIC1(df_weekly)
```

#Comparing MAE values for different models
```{r}
ARModelsMAE <- function(dfAll) {
  statsDF <- data.frame(stations = riverStations,
                              ARMA = numeric(33),
                              ARIMA = numeric(33),
                              AR1 = numeric(33))
  
  for (i in seq_along(riverStations)) {
    dfModel <- dfAll %>% filter(station_id == riverStations[i]) 
  
    model1 <- auto.arima(dfModel$height, d = 0, seasonal = FALSE, stepwise = TRUE, ic = "aic")
    model2 <- auto.arima(dfModel$height, seasonal = FALSE, stepwise = TRUE, ic = "aic")
    model3 <- auto.arima(dfModel$height, start.p = 1, start.q = 0, max.p = 1, max.q = 0)
   
    
    statsDF$ARMA[i] <- accuracy(model1)[,"MAE"]
    statsDF$ARIMA[i] <- accuracy(model2)[,"MAE"]
    statsDF$AR1[i] <- accuracy(model3)[,"MAE"]
  }
  
  return(statsDF)
}

#for daily data
MAEdaily <- ARModelsMAE(df_daily)
#for weekly data
MAEweekly <- ARModelsMAE(df_weekly)
```

#point range plot for MAE: Figure 18
```{r}
MAERangeDF <- data.frame(
                         colour = rep(c("ARMA", "ARIMA", "AR(1)"), each = 2),
                         timescales = factor(c("daily","weekly"), levels = c("daily", "weekly")),
                         median = c(median(MAEdaily$ARMA), median(MAEweekly$ARMA),
                                    median(MAEdaily$ARIMA), median(MAEweekly$ARIMA),
                                    median(MAEdaily$AR1), median(MAEweekly$AR1)), 
                         lower = c(quantile(MAEdaily$ARMA, 0.025), quantile(MAEweekly$ARMA, 0.025),
                                   quantile(MAEdaily$ARIMA, 0.025), quantile(MAEweekly$ARIMA, 0.025),
                                   quantile(MAEdaily$AR1, 0.025), quantile(MAEweekly$AR1, 0.025)),
                         upper = c(quantile(MAEdaily$ARMA, 0.975), quantile(MAEweekly$ARMA, 0.975),
                                   quantile(MAEdaily$ARIMA, 0.975), quantile(MAEweekly$ARIMA, 0.975),
                                   quantile(MAEdaily$AR1, 0.975), quantile(MAEweekly$AR1, 0.975))
                         )


ggplot(MAERangeDF, aes(x = timescales, y = median, color = colour)) +
  theme_minimal() +
  geom_pointrange(aes(ymin = lower, ymax = upper), position = position_dodge(width = 1)) +
  geom_text(aes(label = sprintf("%.4f", median)), position = position_dodge(width = 1), vjust = -2, hjust = 0.5, size = 3.4) +
  geom_text(aes(y = lower, label = sprintf("%.4f", lower)), position = position_dodge(width = 1), vjust = 1, hjust = 0.5, size = 3.4) +
  geom_text(aes(y = upper, label = sprintf("%.4f", upper)), position = position_dodge(width = 1), vjust = -0.4, hjust = 0.5, size = 3.4) +
  xlab("timescale") +
  ylab("MAE (m)") +
  scale_color_manual(values = c("ARMA" = "skyblue", "ARIMA" = "black","AR(1)" = "red"))+
  labs(color = "models") +
  theme(axis.text.x = element_text(size = 10, face = "bold")) +
  theme(axis.text.y = element_text(size = 10, face = "bold")) +
  theme(panel.grid.minor= element_blank(),
        axis.title.x = element_text(size = 12, face = "bold"), 
        axis.title.y = element_text(size = 12, face = "bold"))
```



4.4 State-space models (LL, LLT, LLT with trends)
#AIC values for models of LL, LLT 
```{r}
getAIC <- function(model) {
  return(2*length(model$coef) - 2*model$loglik)
}

SSMAIC <- function(dfAll) {
  statsDF <- data.frame(stations = riverStations,
                              LL = numeric(33),
                              LLT = numeric(33))
  
  for (i in seq_along(riverStations)) {
    dfModel <- dfAll %>% filter(station_id == riverStations[i]) 
  
    model1 <- StructTS(dfModel$height, type = "level")
    model2 <- StructTS(dfModel$height, type = "trend")
    
    statsDF$LL[i] <- getAIC(model1)
    statsDF$LLT[i] <- getAIC(model2)
  }
  
  statsDF$diff <- statsDF$LL - statsDF$LLT
  return(statsDF)
}

#for daily data
SSMAIC(df_daily)
#for weekly data
SSMAIC(df_weekly)
```

#MAE values for models of LL, LLT
```{r}
getMAE <- function(dfSSM, model) {
  fitValue <- as.numeric(fitted(model)[,"level"])
  return(mean(abs(dfSSM$height - fitValue)))
}

SSMMAE <- function(dfAll) {
  statsDF <- data.frame(stations = riverStations,
                              LL = numeric(33),
                              LLT = numeric(33))
  
  for (i in seq_along(riverStations)) {
    dfModel <- dfAll %>% filter(station_id == riverStations[i]) 
  
    model1 <- StructTS(dfModel$height, type = "level")
    model2 <- StructTS(dfModel$height, type = "trend")
    
    statsDF$LL[i] <- getMAE(dfModel, model1)
    statsDF$LLT[i] <- getMAE(dfModel, model2)
  }
  
  return(statsDF)
}

#for daily data
SSMMAE(df_daily)
#for weekly data
SSMMAE(df_weekly)
```

#Getting estimated parameters for LL and LLT
```{r}
#for LL models (level, epsilon)
LLParameterFunction <- function(dfAll) {
  LLpara <- data.frame(stations = riverStations,
                        level = numeric(33),
                        epsilon = numeric(33))
  
  for (i in seq_along(riverStations)) {
    dfLL <- dfAll %>% filter(station_id == riverStations[i])
    LLts <- ts(dfLL$height, start = 1, frequency = 1)
    LLModel <- StructTS(LLts, type = "level")
    
    LLpara$level[i] <- as.numeric(LLModel$coef[1])
    LLpara$epsilon[i] <- as.numeric(LLModel$coef[2])
  }
  
  return(LLpara)
}

#for LLT models (level, slope, epsilon)
LLTParameterFunction <- function(dfAll) {
  LLTpara <- data.frame(stations = riverStations,
                        level = numeric(33),
                        slope = numeric(33),
                        epsilon = numeric(33))
  
  for (i in seq_along(riverStations)) {
    dfLLT <- dfAll %>% filter(station_id == riverStations[i])
    LLTts <- ts(dfLLT$height, start = 1, frequency = 1)
    LLTModel <- StructTS(LLTts, type = "trend")
    
    LLTpara$level[i] <- as.numeric(LLTModel$coef[1])
    LLTpara$slope[i] <- as.numeric(LLTModel$coef[2])
    LLTpara$epsilon[i] <- as.numeric(LLTModel$coef[3])
  }
  
  return(LLTpara)
}

#for daily data
LLParameterFunction(df_daily)
LLTParameterFunction(df_daily)
#for weekly data
LLParameterFunction(df_weekly)
LLTParameterFunction(df_weekly)
```


#time series cross validation MAE for LL, LLT and AR(1)
```{r}
LLCrossValidation <- function(dfAll) {
  crossValidationDF <- data.frame(stations = riverStations,
                                  MAELL = numeric(33),
                                  MAELLT = numeric(33),
                                  MAEAR = numeric(33))
  
  for (i in seq_along(riverStations)) {
    dfLL <- dfAll %>% filter(station_id == riverStations[i])
    LLts <- ts(dfLL$height, start = 1, frequency = 1)
    
    LLForecast <- function(y, h) {
      fit <- StructTS(y, type = "level")
      fc <- forecast(fit, h = h)
      return(fc)
    }
    
    LLTForecast <- function(y, h) {
      fit <- StructTS(y, type = "trend")
      fc <- forecast(fit, h = h)
      return(fc)
    }
    
    ARForecast <- function(y, h) {
      fit <- arima(y, order = c(1,0,0))
      fc <- forecast(fit, h = h)
      return(fc)
    }
    
    errorCVLL <- as.numeric(tsCV(LLts, LLForecast, h = 1))
    errorCVLLT <- as.numeric(tsCV(LLts, LLTForecast, h = 1)) 
    errorCVAR <- as.numeric(tsCV(LLts, ARForecast, h = 1))
    
    crossValidationDF$MAELL[i] = mean(abs(errorCVLL), na.rm = TRUE)
    crossValidationDF$MAELLT[i] = mean(abs(errorCVLLT), na.rm = TRUE)
    crossValidationDF$MAEAR[i] = mean(abs(errorCVAR), na.rm = TRUE)
  }
  
  return(crossValidationDF)
}

#for daily data
LLCV <- LLCrossValidation(df_daily)
LLCV$min <- apply(LLCV[, c("MAELL", "MAELLT", "MAEAR")], 1, function(x) {
  which.min(x)  
})

#for weekly data
LLCV2 <- LLCrossValidation(df_weekly)
LLCV2$min <- apply(LLCV2[, c("MAELL", "MAELLT", "MAEAR")], 1, function(x) {
  which.min(x)  
})
```

#Point range plot for cross validation MAE
```{r}
CVMAERangeDF <- data.frame(
                         colour = rep(c("LL", "LLT", "AR(1)"), each = 2),
                         timescales = factor(c("daily","weekly"), levels = c("daily", "weekly")),
                         median = c(median(LLCV$MAELL), median(LLCV2$MAELL),
                                    median(LLCV$MAELLT), median(LLCV2$MAELLT),
                                    median(LLCV$MAEAR), median(LLCV2$MAEAR)), 
                         lower = c(quantile(LLCV$MAELL, 0.025), quantile(LLCV2$MAELL, 0.025),
                                   quantile(LLCV$MAELLT, 0.025), quantile(LLCV2$MAELLT, 0.025),
                                   quantile(LLCV$MAEAR, 0.025), quantile(LLCV2$MAEAR, 0.025)),
                         upper = c(quantile(LLCV$MAELL, 0.975), quantile(LLCV2$MAELL, 0.975),
                                   quantile(LLCV$MAELLT, 0.975), quantile(LLCV2$MAELLT, 0.975),
                                   quantile(LLCV$MAEAR, 0.975), quantile(LLCV2$MAEAR, 0.975))
                         )


ggplot(CVMAERangeDF, aes(x = timescales, y = median, color = colour)) +
  theme_minimal() +
  geom_pointrange(aes(ymin = lower, ymax = upper), position = position_dodge(width = 1)) +
  geom_text(aes(label = sprintf("%.4f", median)), position = position_dodge(width = 1), vjust = -2, hjust = 0.5, size = 3.4) +
  geom_text(aes(y = lower, label = sprintf("%.4f", lower)), position = position_dodge(width = 1), vjust = 1, hjust = 0.5, size = 3.4) +
  geom_text(aes(y = upper, label = sprintf("%.4f", upper)), position = position_dodge(width = 1), vjust = -0.4, hjust = 0.5, size = 3.4) +
  xlab("timescale") +
  ylab("Cross validation MAE (m)") +
  scale_color_manual(values = c("LL" = "skyblue","LLT" = "black","AR(1)" = "red"))+
  labs(color = "models") +
  theme(axis.text.x = element_text(size = 10, face = "bold")) +
  theme(axis.text.y = element_text(size = 10, face = "bold")) +
  theme(panel.grid.minor= element_blank(),
        axis.title.x = element_text(size = 12, face = "bold"), 
        axis.title.y = element_text(size = 12, face = "bold"))
```


